---
title: Useful data science functions in R
linktitle: R data science functions
toc: true
type: docs
date: "2019-07-17T00:00:00+08:00"
lastmod: "2019-07-17T00:00:00+08:00"
draft: false
menu:
  datascience:
    parent: Notes
    weight: 2

# Prev/next pager order (if `docs_section_pager` enabled in `params.toml`)
weight: 1
output:
  html_document:
    keep_md: yes
---

```{r, include = FALSE}
delete_files <- function() {
  file.types <- c("*.html")
  for (i in seq_along(file.types)) {
    file.remove(dir(pattern = file.types[i], full.names = TRUE))
  }
}
delete_files()
```

As usual, before starting, load all the packages you need.

```{r}
library(tidyverse)
```

# Data management

## Importing data sets

Data can be imported by going to $\text{File}\rightarrow\text{Import Dataset}$.
Alternatively, the code is

```{r}
# code for importing data
```

## Converting to tibbles

```{r}
as_tibble(iris)
```

## Subsetting

To extract columns, use the `$` symbol.

```{r}
iris$Sepal.Length
```

## Reshaping

```{r}
reshape2::melt(table(diamonds$cut, diamonds$color), 
                      var = c("cut", "color"))
```

# Variation

## Continuous variables

### 5-point summaries

```{r}
summary(iris)
ggplot(reshape2::melt(iris), aes(x = variable, y = value)) +
  geom_boxplot()
```

Note: in the above code, we used the `melt()` function in `reshape2` package to aggregate the data. 
Explore what `melt()` does by running it in the console.

### Histograms

```{r}
ggplot(iris, aes(x = Sepal.Length)) +
  geom_histogram()
```

### Density plots

```{r}
ggplot(iris, aes(x = Sepal.Length)) +
  geom_density(fill = "blue")
```

## Discrete variables

### Frequency tables

```{r}
table(mpg$class)
prop.table(table(mpg$class))
```

### Barplots

```{r}
ggplot(mpg, aes(x = reorder(class, class, FUN = length))) +
  geom_bar() +
  labs(x = "Class")
```

Note: The `reorder()` function sorts the bars... the syntax is a bit tricky to understand, so take it as is for now.

# Covariation

## Continuous variables

### Scatter plots

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() 
```

### Smooth lines

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth() 
```

### Binning

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_boxplot(aes(group = cut_width(displ, 0.5)))
```

## Discrete variables

### Contingency tables

```{r}
table(diamonds$cut, diamonds$color)
t(table(diamonds$cut, diamonds$color))
```

### Count plots

```{r}
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
```

### Heat maps

```{r}
dat <- reshape2::melt(table(diamonds$cut, diamonds$color), 
                      var = c("cut", "color"))
ggplot(dat, aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = value))
```

## Continuous x discrete

### Faceting

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_wrap(~ class, nrow = 2)
```

### Adding additional aesthetics

```{r}
ggplot(mpg, aes(x = displ, y = hwy, col = class)) +
  geom_point() 
```
